Publication Bias in Meta-Analysis

### 2019-05-20 Presentation/present.Rnw
Präsentation

dataset.properties <- rbind(study.rev, comparison.rev, meta.rev, study.years, sample.size)
rownames(dataset.properties) <- c("Study number", "Comparison number", "Group number", "Study years", "Study sample size")
colnames(dataset.properties) <- c("5% quantile", "median", "mean", "95% quantile")

@@

@@

dat_text <- data.frame(
  label = paste(c(sum(metac.cont$begg.test), sum(metac.cont$egger.test), sum(metac.cont$thomson.test)), "> 0.05"),
  test.type   = c("pval.begg.cont", "pval.egger.cont", "pval.thomson.cont"))

@@

dat_text <- data.frame(
  label = paste(c(sum(metac.bin$egger.test), sum(metac.bin$harbord.test), sum(metac.bin$peter.test), 
  										sum(metac.bin$rucker.test), sum(metac.bin$schwarzer.test)), "> 0.05"),
  test.type   = c("pval.egger.bin", "pval.harbord.bin", "pval.peter.bin", "pval.rucker.bin", "pval.schwarzer.bin"))

p.dist.bin <- metac.bin %>% ungroup() %>% 
  select(pval.egger.bin, pval.harbord.bin, pval.peter.bin, pval.rucker.bin, pval.schwarzer.bin) %>% 
  gather(key = "test.type", value = "p.value") %>% 
  ggplot(aes(x = p.value)) + geom_histogram(bins = 20) + theme_bw() + facet_wrap(~test.type) + 
	geom_text(data = dat_text, mapping = aes(x = -Inf, y = -Inf, label = label), hjust   = -0.65,
  vjust   = -4, color = "white")

#Test agreement
agree.bin <- meta.bin %>% mutate(n.sig = peter.test + rucker.test + egger.test + harbord.test + schwarzer.test) %>% 
  group_by(n.sig) %>% count %>% filter(n.sig > 0) %>% 
  ggplot(aes(y = nn, x = n.sig)) + theme_bw() + geom_col() + xlab("Number of significant tests") + ylab("count") + ggtitle("Binary Outcomes")

agree.cont <- meta.cont %>% mutate(n.sig = egger.test + thomson.test + begg.test) %>% 
  group_by(n.sig) %>% count %>% filter(n.sig > 0) %>% 
  ggplot(aes(y = nn, x = n.sig)) + theme_bw() + geom_col() + xlab("Number of significant tests") + ylab("count") + ggtitle("Continuous Outcomes")

@@

@

@@

\begin{frame}{Cochrane Library}
Database of high-quality, systematic reviews in clinical science.

Currently $\sim$ 8,000 reviews, prepared by independent groups.

Reviews are peer-reviewed and prepared after guidelines.
\end{frame}

@@

@@

\begin{frame}{Dataset Structure}
\begin{itemize}
\item Comparison: What is compared, e.g. treatment vs. control
\item Outcome: How it is compared
\item Subgroup: Subgroup affiliation
\item Meta-Analysis Group: Results from same comparison, outcome and subgroup
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Review Example: binary outcome}
Barbiturate efficacy for head injury treatment
\vspace{-5mm}
<<echo = FALSE, results = 'asis'>>=
print(xtable(barbi2, label = "barbiturates", digits = 0), include.rownames = F, size = "tiny")
@
\end{frame}

\begin{frame}[fragile]{Dataset Properties}
Missing data:
<<echo=FALSE, results = 'asis'>>=
print(xtable(missing.table, align = "lr"), include.rownames = T, include.colnames = F)
@
\end{frame}

\begin{frame}[fragile]{Dataset Properties}
Review and study properties:
<<echo=FALSE, results = 'asis'>>=
print(xtable(dataset.properties, digits = 0, align = "lrrrr"), include.rownames = T, size = "footnotesize", hline = c(0,1,1,4))
@
\end{frame}

\begin{frame}[fragile]{Pooling Studies - Meta-Analysis}
Multiple results in a meta-analysis group can be pooled:
<<echo=FALSE, results = 'asis'>>=
print(xtable(cum.repr.trials.subg, label = "repr.groups", align = "lrrr", digits = 0), include.rownames = F, size = "footnotesize")
@
\end{frame}

\begin{frame}{Meta-analysis}
Benefits:
\begin{itemize}
\item Summary of evidence (e.g. of a treatment)
\item More reliable evidence (?)
\end{itemize}

@@

\begin{frame}{Small Study Effects}
Causes:
\begin{itemize}
\item Selective publication of studies with large effects
\item Bias in smaller studies
\item Systematical differences in study settings
\item \ldots
\end{itemize}
\end{frame}

\begin{frame}{Small Study Effect Tests}
Tests applicable if:
\begin{itemize}
\item $n$ large
\item variation in the estimated variances of effects \\(here: $\frac{\sigma_{\textrm{max}}}{\sigma_{\textrm{min}}} > 4$)
\end{itemize}
Adjustments required if variance is dependent on effect size (e.g. log odds ratios)
\end{frame}

@@

<<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
par(las = 1, mfrow = c(1,2))
radial(meta.example, comb.fixed = F)
radial(meta.nexample, comb.fixed = F)
@

@@

\begin{frame}{Regression based Tests}
Continuous outcomes tests:

$i$ being the $i$th study in a meta-analysis:

Let $y_{i} = \textrm{effect}_{i}/\textrm{se}_{i}$ and $x_{i} = 1/\textrm{se}_{i}$
\begin{itemize}
\item \citet{Egger} : \\ $y_{i} \sim N(\beta_{0} + \beta_{1}x_{i}, v_{i})$
\item \citet{thompson.sharp} : \\ $y_{i} \sim N(\beta_{0} + \beta_{1}x_{i}, v_{i} + \tau^2)$
\end{itemize}
%Test for non-zero intercept $\beta_{0}$
\end{frame}

\begin{frame}[fragile]{Egger's Tests}

\vspace{-6cm}

<<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
par(las = 1, mfrow = c(1,2))
metabias(meta.example, plotit = TRUE, method.bias = "linreg")
metabias(meta.nexample, plotit = TRUE, method.bias = "linreg")
@

@@

\begin{frame}[fragile]{Thompson and Sharp's Tests}

\vspace{-6.5cm}

<<message = FALSE, echo=FALSE, warning = FALSE, fig.height= 4>>=
par(las = 1, mfrow = c(1,2))
metabias(meta.example, plotit = TRUE, method.bias = "mm")
metabias(meta.nexample, plotit = TRUE, method.bias = "mm")
@

\begin{frame}{Regression based Tests}
Adjustments for binary outcomes:
\begin{itemize}
\item \citet{Peters} :$x_{i}$ = inverse of total sample size, $\textrm{variance}_{i}$ as weight.
\item \citet{Harbord} :$x:{i}$ = score of the log-likelihood of a proportion, $\textrm{variance}_{i}$ as weight.
\item \citet{Rucker} :Use arcsine variance stabilizing transformation for variances and effects, do e.g. Egger's test.
\end{itemize}
\end{frame}

\begin{frame}{Rank based tests}
\citet{begg.ties}
Let $y_{i}$ the standardized effect size of a study $i$, $v_{i}$ it's variance and $n$ the number of studies

$u$ the number of pairs $(y_{i}, v_{i})$ ranked in the same order, $l$
the number of pairs in the opposite order

$Z = (u - l)/\sqrt{n(n-1)(2n + 5)/18}$ \\ can be used as a test statistic (based on Kendall's Tau)
\end{frame}

\begin{frame}{Rank based tests}
\citet{Schwarzer}
Let $e_{t}$ the number of events in the treatment group

Given constant log odds ratio, $E_{t}$ follows a hypergeometric distribution.

$\mathbb{E}(E_{t})$ and the variance can be estimated and used as in 
\citet{begg.ties}
\end{frame}

\begin{frame}{Test Results}
Application of tests if:
\begin{itemize}
\item $n \geq 10$
\item at least one statistically significant effect in a study
\item $\frac{\sigma_{\textrm{max}}}{\sigma_{\textrm{min}}} > 4$
\item $I^2 < 0.5$
\end{itemize}

@@

\begin{frame}[fragile]{Continuous Outcome Test Results}
$p$-values distribution:

<<echo = FALSE, fig.height = 4, message = FALSE>>=
plot(p.dist.cont)
@
\end{frame}

\begin{frame}[fragile]{Binary Outcome Test Results}
$p$-values distribution:

<<echo = FALSE, fig.height = 4, message = FALSE>>=
plot(p.dist.bin)
@
\end{frame}

@@

<<echo = FALSE, fig.height = 3.8, message = FALSE>>=
grid.arrange(agree.cont, agree.bin, ncol = 2)
@
\end{frame}

@@

\begin{frame}{Adjustment by regression}
Similar to the tests, but with unnormalized effect $y_{i}$:

\begin{align}
y_{i} = \beta_{0} + \beta_{1}x_{i}
\end{align}

$\beta_{0}$	 corresponds to $y_{i}$ with $x_{i} = 0$
\end{frame}

\begin{frame}[fragile]{Adjustment by regression}
Linear regression method:

<<echo = FALSE, fig.height = 3>>=
grid.arrange(reg.1, reg.2, ncol = 2)
@
\end{frame}

\begin{frame}[fragile]{Shrinkage Regression}
Extended random effects model:

\vspace{-4mm}
\begin{align}
y_{i} = \beta_{0} + \beta_{1}(\sqrt{v_{i} + \tau^2}) + \epsilon_{i}(\sqrt{s_{i} + \tau^2}),  
\epsilon_{i} \stackrel{\textrm{iid}}{\sim} N(0,1) \nonumber
\end{align}

\vspace{-8mm}
\begin{align}
\mathbb{E}(y_{i}) \rightarrow \beta_{0} + \beta_{1}\tau \textrm{ if } \sqrt{v_{i}} \rightarrow 0 \nonumber
\end{align}

$\beta_{0}, \beta_{1}$ and $\tau$ can be estimated e.g. by ML and REML.
\end{frame}

\begin{frame}[fragile]{Shrinkage Regression}
shrinking the within study variance:

\vspace{-4mm}
\begin{align}
y_{i} &= \beta_{0}^\star + \beta_{1}^\star(\sqrt{v_{i}/M + \tau^2}) + \epsilon_{i}(\sqrt{s_{i}/M + \tau^2}) \nonumber
\end{align}

Letting $M \rightarrow \infty$ and substituting for all parameters and the observed residual

\vspace{-8mm}
\begin{align}
y_{\infty,i} &= \beta_{0}^\star + \sqrt{\frac{\tau^2}{v_{i} + \tau^2}}(y_{i} - \beta_{0}^\star)
\end{align}

\begin{frame}[fragile]{Shrinkage Regression}
Three different treatment effect estimates:

\begin{align}
y_{\infty,i} &= \beta_{0}^\star + \sqrt{\frac{\tau^2}{v_{i} + \tau^2}}(y_{i} - \beta_{0}^\star)
\end{align}

>>>>

dataset.properties <- rbind(study.rev, comparison.rev, meta.rev, study.years, sample.size)
rownames(dataset.properties) <- c("Number of studies", "Number of comparisons", "Number of meta-analyses", "Study years", "Study sample size")
colnames(dataset.properties) <- c("5% quantile", "median", "mean", "95% quantile")

@@

meta.id <- function(meta.id.tolook){
	dt <- data.ext %>% filter(meta.id == meta.id.tolook)
	if(any(is.na(dt$events1)) | all(dt$events1 == 0)){
		meta.ex <- metacont(n.e = total1, mean.e = mean1, sd.e = sd1, n.c = total2, mean.c = mean2, sd.c = sd2, studlab = study.name, data = dt)
	} else{
		meta.ex <- metabin(event.e = events1, n.e = total1, event.c = events2, n.c = total2, studlab = study.name, sm = "RR", data = dt)
	}
	return(meta.ex)
}

@@

dat_text <- data.frame(
  label = paste(c(sum(metac.cont$begg.test), sum(metac.cont$egger.test), sum(metac.cont$thomson.test)), "< 0.05"),
  test.type   = c("pval.begg.cont", "pval.egger.cont", "pval.thomson.cont"))

@@

dat_text <- data.frame(
  label = paste(c(sum(metac.bin$egger.test), sum(metac.bin$harbord.test), sum(metac.bin$peter.test), 
  										sum(metac.bin$rucker.test), sum(metac.bin$schwarzer.test)), "< 0.05"),
  test.type   = c("pval.egger.bin", "pval.harbord.bin", "pval.peter.bin", "pval.rucker.bin", "pval.schwarzer.bin"))

p.dist.bin <- metac.bin %>% ungroup() %>% 
  select(pval.egger.bin, pval.harbord.bin, pval.peter.bin, pval.rucker.bin, pval.schwarzer.bin) %>% 
  gather(key = "test.type", value = "p.value") %>% 
  ggplot(aes(x = p.value)) + geom_histogram(bins = 20) + theme_bw() + facet_wrap(~test.type) + 
	geom_text(data = dat_text, mapping = aes(x = -Inf, y = -Inf, label = label), hjust   = -0.65,
  vjust   = -2, color = "white")

#Test agreement
agree.bin <- metac.bin %>% mutate(n.sig = peter.test + rucker.test + egger.test + harbord.test + schwarzer.test) %>% 
  group_by(n.sig) %>% count %>% filter(n.sig > 0) %>% 
  ggplot(aes(y = nn, x = n.sig)) + theme_bw() + geom_col() + xlab("Number of significant tests") + ylab("count") + ggtitle("Binary Outcomes")

agree.cont <- metac.cont %>% mutate(n.sig = egger.test + thomson.test + begg.test) %>% 
  group_by(n.sig) %>% count %>% filter(n.sig > 0) %>% 
  ggplot(aes(y = nn, x = n.sig)) + theme_bw() + geom_col() + xlab("Number of significant tests") + ylab("count") + ggtitle("Continuous Outcomes")

@@

#Adjustment results: Effect difference histograms
effect.diff <- metac %>% mutate(
	est.fixef = ifelse(!is.na(est.fixef.bin), est.fixef - 1, est.fixef),
	est.ranef = ifelse(!is.na(est.fixef.bin), est.ranef - 1, est.ranef),
	est.trimfill.fixef = ifelse(!is.na(est.fixef.bin), est.trimfill.fixef - 1, est.trimfill.fixef),
	est.copas = ifelse(!is.na(est.fixef.bin), est.copas - 1, est.copas),
	est.reg.ranef = ifelse(!is.na(est.fixef.bin), est.reg.ranef - 1, est.reg.ranef),
	fixef.trimfill = ifelse(abs(est.fixef) > abs(est.trimfill.fixef), "Reduction", "Amplification"),
	fixef.copas = ifelse(abs(est.fixef) > abs(est.copas), "Reduction", "Amplification"),
	fixef.reg = ifelse(abs(est.fixef) > abs(est.reg.ranef), "Reduction", "Amplification"),
	fixef.trimfill.s = ifelse(sign(est.fixef) == sign(est.trimfill.fixef), "Unchanged", "Reversed"),
	fixef.copas.s = ifelse(sign(est.fixef) == sign(est.copas), "Unchanged", "Reversed"),
	fixef.reg.s = ifelse(sign(est.fixef) == sign(est.reg.ranef), "Unchanged", "Reversed"))

trimfill.miss <- effect.diff %>% group_by(fixef.trimfill) %>% mutate(dif = abs(est.fixef - est.trimfill.fixef)) %>% 
	filter(dif > 2) %>% mutate(label = round(dif, 1), type = !is.na(est.fixef.bin)) %>% arrange(label) %>% select(label, fixef.trimfill.s, meta.id, type)

trimfill.label <- data.frame(label = paste("missing:", c(paste(trimfill.miss$label[trimfill.miss$fixef.trimfill == "Reduction"], collapse = ", "),
																												 paste(trimfill.miss$label[trimfill.miss$fixef.trimfill == "Amplification"], collapse = ", "))),
														fixef.trimfill = unique(trimfill.miss$fixef.trimfill), 
														fixef.trimfill.s  = "Reversed")

trimfill.label$label <- as.character(trimfill.label$label)
trimfill.label$label[1] <- sub("2.5,", "2.5, \n   ", trimfill.label$label[1])

hist.trimfill <- effect.diff %>% ggplot(aes(x = abs(est.fixef - est.trimfill.fixef), fill = factor(fixef.trimfill.s))) +
	geom_histogram() + facet_grid(~factor(fixef.trimfill)) + xlim(0, 2)+ xlab("Effect difference") +
	geom_text(data = trimfill.label, mapping = aes(label = label, x = 0, hjust = 0, y = 300), 
						color = "black") + scale_fill_discrete(name="Effect direction") + theme_bw() + 
	theme(legend.position = "bottom", strip.text.x = element_text(size=15))

hist.trimfill1 <- effect.diff %>% ggplot(aes(x = abs(est.fixef - est.trimfill.fixef))) +
	geom_histogram() + facet_grid(~factor(fixef.trimfill)) + xlim(0, 2)+ xlab("Effect difference") +
	geom_text(data = trimfill.label, mapping = aes(label = label, x = 0, hjust = 0.05, y = 300), 
						color = "black") + theme_bw() + 
	theme(legend.position = "bottom", strip.text.x = element_text(size=15))

funnel(trimfill(meta.id(165815)))

copas.miss <- effect.diff %>% filter(!is.na(est.copas)) %>% group_by(fixef.copas) %>% mutate(dif = abs(est.fixef - est.copas)) %>% 
	filter(dif > 2) %>% mutate(label = round(dif, 1), type = !is.na(est.fixef.bin)) %>% arrange(label) %>% select(label, fixef.copas.s, meta.id, type)

copas.label <- data.frame(label = paste("missing:", c(paste(copas.miss$label[copas.miss$fixef.copas == "Reduction"], collapse = ", "),
																												 paste(copas.miss$label[copas.miss$fixef.copas == "Amplification"], collapse = ", "))),
														 fixef.copas = unique(copas.miss$fixef.copas), 
														 fixef.copas.s  = "Reversed")

copas.label$label <- as.character(copas.label$label)
copas.label$label[1] <- sub("3.5,", "3.5, \n   ", copas.label$label[1])

hist.copas <- effect.diff %>% filter(!is.na(est.copas)) %>% ggplot(aes(x = abs(est.fixef - est.copas), fill = factor(fixef.copas.s))) +
	geom_histogram() + facet_grid(~factor(fixef.copas)) + xlim(0, 2) + xlab("Effect difference") +
	geom_text(data = copas.label, mapping = aes(label = label, x = 0, hjust = 0, y = 175), 
						color = "black") + scale_fill_discrete(name="Effect direction") + theme_bw() +
	theme(legend.position = "bottom", strip.text.x = element_text(size=15))

hist.copas1 <- effect.diff %>% filter(!is.na(est.copas)) %>% ggplot(aes(x = abs(est.fixef - est.copas))) +
	geom_histogram() + facet_grid(~factor(fixef.copas)) + xlim(0, 2) + xlab("Effect difference") +
	geom_text(data = copas.label, mapping = aes(label = label, x = 0, hjust = 0.05, y = 175), 
						color = "black") + theme_bw() +
	theme(legend.position = "bottom", strip.text.x = element_text(size=15))

funnel(meta.id(56200))

reg.ranef.miss <- effect.diff %>% group_by(fixef.reg) %>% mutate(dif = abs(est.fixef - est.reg.ranef)) %>% 
	filter(dif > 3) %>% mutate(label = round(dif, 1), type = !is.na(est.fixef.bin)) %>% arrange(label) %>% select(label, fixef.reg.s, meta.id, type)

reg.ranef.label <- data.frame(label = paste("missing:", c(paste(reg.ranef.miss$label[reg.ranef.miss$fixef.reg == "Reduction"], collapse = ", "),
																											paste(reg.ranef.miss$label[reg.ranef.miss$fixef.reg == "Amplification"], collapse = ", "))),
													fixef.reg = unique(reg.ranef.miss$fixef.reg), 
													fixef.reg.s  = "Reversed")

reg.ranef.label$label <- as.character(reg.ranef.label$label)
reg.ranef.label$label[1] <- sub("4.8,", "4.8,\n   ", reg.ranef.label$label[1])

hist.reg <- effect.diff %>% ggplot(aes(x = abs(est.fixef - est.reg.ranef), fill = factor(fixef.reg.s))) +
	geom_histogram() + facet_grid(~factor(fixef.reg)) + xlim(0, 3)+ xlab("Effect difference") +
	geom_text(data = reg.ranef.label, mapping = aes(label = label, x= 0, y = 250, hjust = 0.05), 
						color = "black", size = 2.8) + theme_bw() + scale_fill_discrete(name="Effect direction") +
	theme(legend.position = "bottom", strip.text.x = element_text(size=15))

hist.reg1 <- effect.diff %>% ggplot(aes(x = abs(est.fixef - est.reg.ranef))) +
	geom_histogram() + facet_grid(~factor(fixef.reg)) + xlim(0, 3)+ xlab("Effect difference") +
	geom_text(data = reg.ranef.label, mapping = aes(label = label, x= 0, y = 250, hjust = 0.05), 
						color = "black", size = 2.8) + theme_bw() + 
	theme(legend.position = "bottom", strip.text.x = element_text(size=15))

metac %>% filter(meta.id == 101436) %>% select(comparison.name, outcome.name, sungroup.name, est.fixef, est.copas, 
																							 est.trimfill.fixef, est.reg.ranef, est.copas)
par(mfrow = c(1,3))
funnel(meta.id(101436))
funnel(trimfill(meta.id(101436)))
funnel(limitmeta(meta.id(101436)))

print(effect.diff %>% group_by(fixef.reg) %>% mutate(dif = abs(est.fixef - est.reg.ranef)) %>% 
				filter(dif > 3) %>% mutate(label = round(dif, 1), type = !is.na(est.fixef.bin)) %>% 
				arrange(label) %>% select(label, type, meta.id, comparison.name, outcome.name), n = 200)

#Scatterplots of metac-analysis and corrected log(abs(abs estimates
trimfill.fixef.sc.effect <- metac %>% ggplot(aes(x = log(abs((est.fixef))), y = log(abs(est.trimfill.fixef)))) + geom_point() + 
	geom_abline(slope = 1, color = "red") + geom_smooth(method = "lm", se = FALSE) + 
	xlab("Fixed effects log estimate") + ylab("Trimfill log estimate") + ggtitle("Fixed effects to Trimfill")

trimfill.ranef.sc.effect <- metac %>% ggplot(aes(x = log(abs((est.ranef))), y = log(abs(est.trimfill.fixef)))) + geom_point() + 
	geom_abline(slope = 1, color = "red") + geom_smooth(method = "lm", se = FALSE) + 
	xlab("Random effects log estimate") + ylab("Trimfill log estimate") + ggtitle("Random effects to Trimfill")

#Scatterplots of metac-analysis and corrected log estimates
copas.fixef.sc.effect <- metac %>% ggplot(aes(x = log(abs((est.fixef))), y = log(abs(est.copas)))) + geom_point() + 
	geom_abline(slope = 1, color = "red") + geom_smooth(method = "lm", se = FALSE) + 
	xlab("Fixed effects log estimate") + ylab("Copas log estimate") + ggtitle("Fixed effects to Copas")

copas.ranef.sc.effect <- metac %>% ggplot(aes(x = log(abs((est.ranef))), y = log(abs(est.copas)))) + geom_point() + 
	geom_abline(slope = 1, color = "red") + geom_smooth(method = "lm", se = FALSE, show.legend = T) + 
	xlab("Random effects log estimate") + ylab("Copas log estimate") + ggtitle("Random effects to Copas")

#Scatterplots of metac-analysis and corrected log estimates
reg.fixef.sc.effect <- metac %>% ggplot(aes(x = log(abs(log(abs((est.fixef))))), y = log(abs(log(abs(est.reg.ranef)))))) + geom_point() + 
	geom_abline(slope = 1, color = "red") + geom_smooth(method = "lm", se = FALSE) + 
	xlab("Fixed effects log estimate") + ylab("Regression log estimate") + ggtitle("Fixed effects to Regression")

reg.ranef.sc.effect <- metac %>% ggplot(aes(x = log(abs((est.ranef))), y = log(abs(est.reg.ranef)))) + geom_point() + 
	geom_abline(slope = 1, color = "red") + geom_smooth(method = "lm", se = FALSE) + 
	xlab("Random effects log estimate") + ylab("Regression log estimate") + ggtitle("Random effects to Regression")

#Scatterplots of test- statistics:
trimfill.fixef.sc.zval <- metac %>% ggplot(aes(x = abs(zval.fixef), y = abs(zval.trimfill.fixef))) + geom_point() +
	geom_abline(slope = 1, color = "red") + geom_smooth(method = "lm", se = FALSE) + theme_bw() + 
	ggtitle("Fixed effects and trimfill z statistics") + ylab("Trimfill  statistic") + xlab("Fixed effects  statistic")

trimfill.ranef.sc.zval <- metac %>% ggplot(aes(x = abs(zval.ranef), y = abs(zval.trimfill.fixef))) + geom_point() +
	geom_abline(slope = 1, color = "red") + geom_smooth(method = "lm", se = FALSE) + theme_bw() + 
	ggtitle("Random effects and trimfill z statistics") + ylab("Trimfill  statistic") + xlab("Random effects  statistic")

#Scatterplots of test- statistics:
copas.fixef.sc.zval <- metac %>% ggplot(aes(x = abs(zval.fixef), y = abs(zval.copas))) + geom_point() +
	geom_abline(slope = 1, color = "red") + geom_smooth(method = "lm", se = FALSE) + theme_bw() + 
	ggtitle("Fixed effects and Copas z statistics") + ylab("Copas  statistic") + xlab("Fixed effects  statistic")

copas.ranef.sc.zval <- metac %>% ggplot(aes(x = abs(zval.ranef), y = abs(zval.copas))) + geom_point() +
	geom_abline(slope = 1, color = "red") + geom_smooth(method = "lm", se = FALSE) + theme_bw() + 
	ggtitle("Random effects and Copas z statistics") + ylab("Copas  statistic") + xlab("Random effects  statistic")

#Scatterplots of test- statistics:
reg.fixef.sc.zval <- metac %>% ggplot(aes(x = abs(zval.fixef), y = abs(zval.reg.ranef))) + geom_point() +
	geom_abline(slope = 1, color = "red") + geom_smooth(method = "lm", se = FALSE) + theme_bw() + 
	ggtitle("Fixed effects and Regression z statistics") + ylab("Regression  statistic") + xlab("Fixed effects  statistic")

reg.ranef.sc.zval <- metac %>% ggplot(aes(x = abs(zval.ranef), y = abs(zval.reg.ranef))) + geom_point() +
	geom_abline(slope = 1, color = "red") + geom_smooth(method = "lm", se = FALSE) + theme_bw() + 
	ggtitle("Random effects and Regression z statistics") + ylab("Regression  statistic") + xlab("Random effects  statistic")
@

@@

% \begin{frame}{Evidence Synthesis}
% Summarise evidence over fields and studies
% 
% Quantitatively by meta-analysis
% 
% \end{frame}

\begin{frame}{Cochrane Organisation}
Aim: summarise findings in primary clinical research and health care

Provide peer-reviewed, systematic reviews

Public access (for some countries)
\end{frame}

@@

\begin{frame}[fragile]{Review Example: Binary Outcome}
Barbiturate efficacy for head injury treatment
\vspace{-5mm}
<<echo = FALSE, results = 'asis'>>=
print(xtable(barbi2, label = "barbiturates", digits = 0), include.rownames = F, size = "tiny")
@
\end{frame}

@@

\begin{frame}[fragile]{Dataset Properties}
Review or study level:
<<echo=FALSE, results = 'asis'>>=
print(xtable(dataset.properties, digits = 0, align = "lrrrr"), include.rownames = T, size = "footnotesize", hline = c(0,1,1,3))
@
\end{frame}

\begin{frame}{Meta-analysis}
Benefits:
\begin{itemize}
\item Summary of evidence (e.g. of a treatment effect)
\item More reliable evidence (?)
\end{itemize}

@@

\begin{frame}{Small Study Effects}
Causes:
\begin{itemize}
\item Selective publication of studies with significant results - publication bias
\item Selective reporting of most favorable outcomes
\item Systematical differences in study settings
\end{itemize}
\end{frame}

\begin{frame}{Small Study Effect Tests}
Tests applicable if:
\begin{itemize}
\item Number of studies $n$ large
\item variation in the estimated variances of effects %\\(here: $\frac{\sigma_{\textrm{max}}}{\sigma_{\textrm{min}}} > 4$)
\end{itemize}
Adjustments required if variance is dependent on effect size (e.g. log odds ratios)
\end{frame}

@@

\vspace{-1.1cm}

<<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
par(las = 1, mfrow = c(1,2))
biased.rev$inv <- 1/biased.rev$se
plot(y = (biased.rev$effect/biased.rev$se), x = (biased.rev$se^-1), xlim = c(0, 6), xlab = "inverse standard error", 
		 ylab = "mean diff. / std. error")

unbiased.rev$inv <- 1/unbiased.rev$se
plot(y = (unbiased.rev$effect/unbiased.rev$se), x = (unbiased.rev$inv), xlim = c(0, 1.2), 
		 ylim = c(-18, 0), xlab = "inverse standard error", ylab = "mean diff. / std. error")
@

@@

\begin{frame}{Regression based Tests}
studies $i,.. ,n$, effects $\theta_i$ and variances $v_i$, s.e. $s_i$

$\theta_M$ is the pooled effect and $\tau^2$ the between-study variance.

Let $y_{i} = \theta_{i}/_{i}$ and $x_i = 1/s_i$
\begin{itemize}
\item \citet{Egger} : Simple linear regression \\ $y_i = \beta_0 + \beta_1 x_i, \epsilon_i \sim N(0, \sigma)$
\item \citet{thompson.sharp} : extension of Egger with study weights $v_{i} + \tau^2$ 
\end{itemize}

@@

\begin{frame}[fragile]{Egger's Test examples}
Test for non-zero intercept $\beta_{0}$

\vspace{-1.1cm}
<<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
par(las = 1, mfrow = c(1,2))
biased.rev$inv <- 1/biased.rev$se
m.ex <- lm(formula = (biased.rev$effect/biased.rev$se) ~ (biased.rev$inv))
plot(y = (biased.rev$effect/biased.rev$se), x = (biased.rev$se^-1), xlim = c(0, 6), xlab = "inverse standard error", 
		 ylab = "mean diff. / std. error")
abline(coef = coef(m.ex), lty = 2, col = 2)

unbiased.rev$inv <- 1/unbiased.rev$se
m.nex <- lm(formula = (unbiased.rev$effect/unbiased.rev$se) ~ (unbiased.rev$inv))
plot(y = (unbiased.rev$effect/unbiased.rev$se), x = (unbiased.rev$inv), xlim = c(0, 1.2), 
		 ylim = c(-18, 0), xlab = "inverse standard error", ylab = "mean diff. / std. error")
abline(coef = coef(m.nex), lty = 2, col = 2)
@
\end{frame}

\begin{frame}{Regression Tests for Binary Outcomes}

\begin{itemize}
\item \citet{Peters} :$x_i = 1/n_i$ instead $1/s_i$, inverse variances as weight.
\item \citet{Harbord} :$x_i$ = score of the log-likelihood of a proportion and inverse variances as weights.
\item \citet{Rucker} :Use arcsine variance stabilizing transformation for variances and effects, do e.g. Egger's test.
\end{itemize}
\end{frame}

\begin{frame}{Rank based tests}
\citet{begg.ties}: \\
Let $y_{i}$ be $frac{\theta_i - \theta_M}{v_i}$ and $x_i$ its variance ($\neq v_i$)

$u$ the number of pairs $(y_{i}, x_{i})$ ranked in the same order, $l$
the number of pairs in the opposite order

$Z = \frac{(u - l)}{\sqrt{n(n-1)(2n + 5)/18}}$  is a test statistic
\end{frame}

\begin{frame}{Rank based tests}
\citet{Schwarzer}: \\
$e_t$ number of events in the treatment group

$E_t$ follows hypergeometric distribution: calculate $\mathbb{E}(E_{t})$ and variances

proceed as in \citet{begg.ties}
\end{frame}

\begin{frame}{Test Results}
Inclusion criteria (from \citet{Ioannidis2007}):
\begin{itemize}
\item $n \geq 10$
\item at least one statistically significant effect in a study
\item $\frac{\sigma_{\textrm{max}}^2}{\sigma_{\textrm{min}}^2} > 4$
\item $I^2 < 0.5$
\end{itemize}

@@

\begin{frame}[fragile]{Continuous Outcome Test Results}
$p$-values distribution, $n$ = \Sexpr{dim(metac.cont)[1]}:

\vspace{-2mm}
<<echo = FALSE, fig.height = 4, message = FALSE>>=
plot(p.dist.cont)
@
\end{frame}

\begin{frame}[fragile]{Binary Outcome Test Results}
$p$-values distribution, $n$ = \Sexpr{dim(metac.bin)[1]}:

\vspace{-2mm}
<<echo = FALSE, fig.height = 3.6, message = FALSE>>=
plot(p.dist.bin)
@
\end{frame}

@@

\vspace{-2mm}
<<echo = FALSE, fig.height = 3.2, message = FALSE>>=
grid.arrange(agree.cont, agree.bin, ncol = 2)
@
\end{frame}

@@

\begin{frame}{Adjustment by regression}
$y_i = \theta_i/s_i, x_i = 1/s_i$

$y_i = \beta_0 + \beta_1 x_i, \epsilon_i \sim N(0, \sigma)$

$\beta_{1}$ is the weighted mean treatment effect if $\beta_{0} = 0$
\end{frame}

\begin{frame}{Adjustment by regression}
Radial plots (continuous outcome examples):

\vspace{-1.1cm}

<<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
par(las = 1, mfrow = c(1,2))
biased.rev$inv <- 1/biased.rev$se
m.ex.0 <- lm(formula = (biased.rev$effect/biased.rev$se) ~ 0 + (biased.rev$inv))
m.ex <- lm(formula = (biased.rev$effect/biased.rev$se) ~ (biased.rev$inv))
plot(y = (biased.rev$effect/biased.rev$se), x = (biased.rev$se^-1), xlim = c(0, 6), xlab = "inverse standard error", 
		 ylab = "mean diff. / std. error")
abline(coef = c(0,coef(m.ex.0)), lty = 2)
abline(coef = coef(m.ex), lty = 2, col = 2)

unbiased.rev$inv <- 1/unbiased.rev$se
m.nex.0 <- lm(formula = (unbiased.rev$effect/unbiased.rev$se) ~ 0 + (unbiased.rev$inv))
m.nex <- lm(formula = (unbiased.rev$effect/unbiased.rev$se) ~ (unbiased.rev$inv))
plot(y = (unbiased.rev$effect/unbiased.rev$se), x = (unbiased.rev$inv), xlim = c(0, 1), 
		 ylim = c(-18, 0), xlab = "inverse standard error", ylab = "mean diff. / std. error")
abline(coef = c(0,coef(m.nex.0)), lty = 2)
abline(coef = coef(m.nex), lty = 2, col = 2)
@
\end{frame}

\begin{frame}[fragile]{Limit Meta-Analysis}
Extended random effects model:

\vspace{-4mm}
\begin{align}
y_{i} = \beta_{0} + \beta_{1}(\sqrt{v_{i} + \tau^2}) + \epsilon_{i}(\sqrt{v_{i} + \tau^2}), \nonumber \\
\epsilon_{i} \stackrel{\textrm{iid}}{\sim} N(0,1) \nonumber
\end{align}

Use $\mathbb{E}(y_{i}) \rightarrow \beta_{0} + \beta_{1}\tau$ for $\sqrt{v_{i}} \rightarrow 0$
as corrected treatment efffect.
\end{frame}

\begin{frame}[fragile]{Shrinkage Regression}
Funnel plot with effect with infinite precision:

\vspace{-1.1cm}
<<message = FALSE, echo=FALSE, warning = FALSE, fig.height=4>>=
par(las = 1, mfrow = c(1,2))
funnel(limitmeta(meta.example))
funnel(limitmeta(meta.nexample))
@
\end{frame}

\begin{frame}{Copas selection model}
\citet{Copas1}: Bivariate normal model with two parts:
